---
title: "Project 2"
format: html
editor: visual
editor_options: 
  chunk_output_type: console
---

## Purpose and Preliminaries

The purpose of this Quarto document is to prepare and explore the data in a static manner that will be used to construct our Shiny app for project 2. We are using the Mobile Device Usage data set. We start by importing the data, as well as packages we will use throughout this Quarto document.

```{r}
library(tidyverse)

user_data <- read_csv('project2-app/user_behavior_dataset.csv')
```

## Preparing the Data

Our first step is to ensure each variable in the tibble is of the appropriate data type. We view the tibble.

```{r}
user_data
```

It appears that Device Model, Operating System, Gender, and User Behavior Class are our categorical variables, whereas the remaining variables are numeric/quantitative. We change the categorical variables so they are stored as factors in our tibble.

```{r}
user_data <- user_data |>
  mutate(`Device Model` = as.factor(`Device Model`),
         `Operating System` = as.factor(`Operating System`),
         Gender = as.factor(Gender),
         `User Behavior Class` = as.factor(`User Behavior Class`))
```

Now that are data are stored in the desired form, we perform our final step of modifying the variable names (when applicable) so that they follow R naming conventions.

```{r}
user_data <- user_data |>
  rename('user_id' = `User ID`,
         'device_model' = `Device Model`,
         'operating_system' = `Operating System`,
         'app_use_time' = `App Usage Time (min/day)`,
         'screen_on_time' = `Screen On Time (hours/day)`,
         'battery_drain' = `Battery Drain (mAh/day)`,
         'app_number' = `Number of Apps Installed`,
         'data_usage' = `Data Usage (MB/day)`,
         'behavior_class' = `User Behavior Class`)
```

## Numerical Summaries: Categorical Variables

We perform some numerical summaries for categorical variables. We use the table function from base R for this purpose. We first produce one-way contingency tables for each categorical variable.

```{r}
table(user_data$device_model)
table(user_data$operating_system)
table(user_data$Gender)
table(user_data$behavior_class)
```

Next we will create two-way contingency tables. We choose a couple of representative pairs of categorical variables for this purpose.

```{r}
user_data |>
  select(device_model, Gender) |>
  table()

user_data |>
  select(behavior_class, operating_system) |>
  table()

```

## Numerical Summaries: Quantitative Variables

Next we perform numerical summaries for quantitative variables at different levels of categorical variables.

Our first example will be summaries of app use time by gender.

```{r}
user_data |>
  group_by(Gender) |>
  summarize('mean' = mean(app_use_time), 
            'median' = median(app_use_time),
            'SD' = sd(app_use_time),
            'IQR' = IQR(app_use_time),
            'minimum' = min(app_use_time),
            'maximum' = max(app_use_time))
```

Our next example will be summaries of app use time by gender and operating system.

```{r}
user_data |>
  group_by(Gender, operating_system) |>
  summarize('mean' = mean(app_use_time), 
            'median' = median(app_use_time),
            'SD' = sd(app_use_time),
            'IQR' = IQR(app_use_time),
            'minimum' = min(app_use_time),
            'maximum' = max(app_use_time))
```

Finally, we will look at number of apps installed by behavior class.

```{r}
user_data |>
  group_by(behavior_class) |>
  summarize('mean' = mean(app_number), 
            'median' = median(app_number),
            'SD' = sd(app_number),
            'IQR' = IQR(app_number),
            'minimum' = min(app_number),
            'maximum' = max(app_number))
```

## Exploratory Data Analysis Plotting

In this section we create several plots to explore our data.

Our first plot will be a scatter plot of app usage time versus number of apps, grouped by gender.

```{r}
user_data |>
  ggplot(aes(app_number, app_use_time)) + geom_point(aes(color = Gender)) + labs(x = 'Number of Apps Installed', y = 'App Usage Time (min/day)', title = 'App Usage Time by Number of Apps Installed and Gender')
```

Our next plot will be a bar plot of behavior class by operating system and gender.

```{r}
user_data |>
  ggplot(aes(behavior_class)) + geom_bar(aes(fill = operating_system), position = 'dodge') + facet_wrap(~ Gender) + labs(x = 'User Behavior Class', title = 'User Behavior Class Counts by Operating System and Gender', fill = 'Operating System')
```

Next we will create a density plot of app use time by gender and device model.

```{r}
user_data |>
  ggplot(aes(app_use_time)) + geom_density(aes(fill = Gender), alpha = 0.5) + facet_wrap(~device_model) + labs(x = 'App Usage Time (min/day)', y = 'Density', title = 'Distribution of App Usage Time by Gender and Device Model') 
```

Next we create a box plot of app number by user behavior class and operating system.

```{r}
user_data |>
  ggplot(aes(x = behavior_class, y = app_number)) + geom_boxplot() + facet_wrap(~operating_system) + labs(x = 'User Behavior Class', y = 'Number of Apps Installed', title = 'Number of Apps Installed by User Behavior Class') + theme_light()
```

Next we create box plots of app use time by gender and operating system.

```{r}
user_data |>
  ggplot(aes(x = Gender, y = app_use_time)) + geom_boxplot() + facet_wrap(~operating_system) + labs(x = 'Gender', y = 'App Usage Time (min/day)', title = 'App Usage Time by Gender and Operating System') + theme_light()
```

Finally we create a plot type that was not covered in class: a correlation matrix. We use the `ggcorrplot` to create the correlation matrix. We round all correlation values to one decimal place.

```{r}
library(ggcorrplot)

corr_matrix <- user_data[unlist(lapply(user_data, is.numeric))] |>
  select(!user_id) |>
  cor() |>
  round(1)

ggcorrplot(corr_matrix, lab = T, title = 'Correlation Matrix for Mobile Device Usage Data', colors = c('yellow', 'blue', 'green'))
  
```

We see from the plot that all of the quantitative variables in the Mobile Device Usage data set are highly positively correlated with one another, except for age, which has nearly no correlation with any of the other variables.
